\section{Performance Evaluation and Conclusion}

\subsection{Hash Algorithm Performance}

All four hash algorithms process messages in constant time relative to input length. The key differentiator is the \textbf{output size}, which determines the signature space and collision resistance:

\begin{table}[H]
\centering
\begin{tabular}{|l|c|c|}
\hline
\textbf{Algorithm} & \textbf{Digest Size} & \textbf{Collision Resistance} \\
\hline
MD5     & 128 bits & $2^{64}$ (broken) \\
SHA-1   & 160 bits & $2^{80}$ (weakened) \\
SHA-256 & 256 bits & $2^{128}$ \\
SHA-512 & 512 bits & $2^{256}$ \\
\hline
\end{tabular}
\caption{Digest sizes and theoretical collision resistance levels.}
\label{tab:performance}
\end{table}

\subsection{Key Size Impact}

The signing and verification operations use Python's built-in \texttt{pow(base, exp, mod)} which implements fast modular exponentiation. With small demonstration primes ($\sim$34-bit keys), both operations complete in microseconds. For production 2048-bit keys, signing typically takes 1–5 ms and verification (with $e = 65537$) completes in under 0.5 ms due to the small Hamming weight of the public exponent.

\subsection{Conclusion}

This project successfully demonstrated the complete RSA digital signature lifecycle through a from-scratch Python implementation. The key findings are:

\begin{itemize}
    \item \textbf{Authenticity:} Only the holder of the private key $d$ can produce a valid signature for a given message.
    \item \textbf{Data Integrity:} Any modification to the signed message is immediately detected through hash comparison, as demonstrated by the tampering attack simulation.
    \item \textbf{Non-repudiation:} The mathematical binding between the signature and the private key prevents the signer from denying their involvement.
\end{itemize}

\textbf{Future work} could extend this implementation by adopting RSA-PSS padding for production-grade security, using cryptographically secure prime generation for larger key sizes, and integrating X.509 certificate chains for key distribution and trust management.
